Synthesizing Quotient Lenses

### 2017-10-06 papers/QuotientLensSynth/main.tex
Started working on synthesis section

\section{Normal Forms QRE Lenses}
\sam{Transition needed?}
\begin{theorem}\label{normal form}
If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}
\end{theorem}
\begin{proof}
Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
derivation $q : c \Leftrightarrow c'$.
\begin{enumerate}
  \item
  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
  $\ell :
  R \Leftrightarrow S$. Then
  \begin{align*}
  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
  = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
  \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
  \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
  \llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
  \end{align*}
  \item
  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
  \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
\begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ \canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exists a bijective lens $\ell :
  K(c') \Leftrightarrow K(c'')$ such that 
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonize(c'')
\end{align*}
Consequently
\begin{align*}
  \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
  \canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
  (\canonizer(c' \circ c))\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
  Canonize(c'')
  \end{align*}

@@

\section{Synthesizing QRE Lenses}
\label{synth}

By \cref{normal form}, if there is a derivation $q : c \Leftrightarrow c'$ in
$\mathcal{Q}$ then there exists a bijective lens $\ell : K(c) \Leftrightarrow
K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}
In other words, every quotient lens in $\mathcal{Q}$ is the same as an ordinary
bijective lens with canonizers at the ends.\\
Therefore, in order to synthesize a quotient lens $q: c \Leftrightarrow c'$
where $c, c'$ are QREs, it suffices to synthesize a bijective lens $\ell : K(c)
\Leftrightarrow K(c')$.\\
\textcolor{red}{[Citation Needed!]} has already shown that if there
exists a derivation $\ell : R \Leftrightarrow S$, then there exist
regular expressions $R', S'$ and a synthesizable \\
\textcolor{red}{[Explain and make explicit the meaning
of synthesizable]} \\
bijective lens $\ell' :
R' \Leftrightarrow S'$ such that $\mathcal{L}(R) = \mathcal{L}(R')$,
$\mathcal{L}(S) = \mathcal{L}(S')$ and $\llbracket \ell \rrbracket =
\llbracket \ell' \rrbracket$. This immediately implies that if there is a
derivation $q : c \Leftrightarrow c'$ as well as a derivation $\ell : K(c)
\Leftrightarrow K(c')$, then $q$ is synthesizable.
\textcolor{red}{[More Needed?]}
\section{Implementation and Evaluation}
\label{impl}

>>>>

\subsection{Normal Forms QRE Lenses}
Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
\Leftrightarrow c'$ be such that 
\begin{align*}
\llbracket q \rrbracket.get &= \ell \circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \ell^{-1} \circ
\canonizer(c')
\end{align*}
for some bijective lens $\ell$. We now provide a proof that all QRE lenses are
of this form.
\begin{theorem}\label{normal form}
If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}
\end{theorem}
\begin{proof}
Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
derivation $q : c \Leftrightarrow c'$.
\begin{enumerate}
  \item
  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
  $\ell :
  R \Leftrightarrow S$. Then
  \begin{align*}
  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
  = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
  \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
  \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
  \llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
  \end{align*}
  \item
  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
  \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
\begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ \canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exists a bijective lens $\ell :
  K(c') \Leftrightarrow K(c'')$ such that 
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonize(c'')
\end{align*}
Consequently
\begin{align*}
  \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
  \canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
  (\canonizer(c' \circ c))\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
  Canonize(c'')
  \end{align*}

@@

\section{Synthesizing QRE Lenses}
\label{synth}
By Theorem~\ref{normal form}, if there is a derivation $q : c \Leftrightarrow
c'$ of a QRE lens, then there exists a bijective lens $\ell : K(c)
\Leftrightarrow K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}

Now suppose that we wish to synthesize a quotient lens $q$ of type $q: c
\Leftrightarrow c'$ that maps a specified set of input strings to a
specified set of output strings and vice versa.

For example, in the \textsc{Bib}\TeX\ to EndNote problem, the programmer wishes
to synthesize a QRE lens $q$ which maps the QRE
$$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
\re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$ 
to the QRE
$$\perm{\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},
\re{YEAR'}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$
(resp. $\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},$ and
$\re{YEAR'}$) are regular expressions describing the respective
\textsc{Bib}\TeX\; (resp. EndNote) fields of the same name, and
$\re{WHITESPACE}$ is a regular expression describing whitespace
characters. Morever, the synthesized lens should map the equivalence
class of this \textsc{Bib}\TeX\; record
\begin{verbatim}
@Book {conway,
  Author = {Conway, J. H.},
  Title = {Regular Algebra and Finite Machines},
  Publisher = {Printed in GB by William Clowes & Sons Ltd},
  Year = {1971}
}
\end{verbatim}
\noindent to the equivalence class of this EndNote entry:
\begin{verbatim}
%0 Book
%T Regular Algebra and Finite Machines
%A Conway, J. H.
%D 1971
%I Printed in GB by William Clowes & Sons Ltd
%F conway
\end{verbatim}

In our previous work, we showed that given regular expressions $R, S$ and a set
of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if there is a
derivation $\ell :, R \Leftrightarrow S$ of a bijective lens that maps $r_i$ to
$s_i$ in the forward direction and $s_i$ to $r_i$ in the backward direction,
then there is an algorithm that computes a bijective lens $\ell' : R'
\Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps $r_i$
to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
direction, and $\llbracket \ell \rrbracket = \llbracket \ell'
\rrbracket$~\cite{popl18}.

\section{Implementation and Evaluation}
\label{impl}
